---
title: "Network Analysis Visualizations"
author: "Aaron Kessler"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    code_folding: show
    df_print: kable
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(kableExtra)
library(here)
library(janitor)
library(igraph)
library(tidygraph)
library(ggraph)
options(dplyr.summarise.inform = FALSE)

#load dataset
top_ten_consolidated_casecounts <- readRDS("data/table_data/top_ten_consolidated_casecounts.rds")


```

# Social Network Analysis

Let's take a look at our data

```{r}

head(top_ten_consolidated_casecounts)

```

## Create the node and edge network datasets with igraph

Edges

```{r}

edges <- top_ten_consolidated_casecounts %>% 
  select(source = clean_firm_consolidated,
         target = parent_company, 
         weight = cases)

edges %>% head(10)

```

Nodes

```{r}
names1 <- top_ten_consolidated_casecounts %>% 
  ungroup() %>% 
  select(name = parent_company) %>% 
  distinct() %>% 
  mutate(group = "company")

head(names1)
```

```{r}
names2 <- top_ten_consolidated_casecounts %>% 
  ungroup() %>% 
  select(name = clean_firm_consolidated) %>% 
  distinct() %>% 
  mutate(group = "lawfirm")

head(names2)
```

```{r}
nodes <- bind_rows(names1, names2)

head(nodes)
```

Add id columns with sequential values.\
Will zero-index the values to work with igraph more smoothly.

```{r}

nodes <- tibble::rowid_to_column(nodes, "id") %>% 
  mutate(id = id - 1) %>% 
  select(name, id, group)

nodes %>% 
  head(10)

```

### igraph object creation

```{r}

g <- graph_from_data_frame(d = edges, 
                           vertices = nodes, 
                           directed = FALSE)
g


```

Access nodes

```{r}

V(g)


```

Names of each node

```{r}

head(V(g)$name)


```

Edges

```{r}

E(g) # edges


```

Weights for each edge

```{r}

E(g)$weight # weights for each edge


```

attributes of edges

```{r}

edge_attr(g) 


```

# Network visualization

## Using Tidygraph for node and edge sets

Create the tbl_graph object.

```{r}
# https://www.jessesadler.com/post/network-analysis-with-r/

routes_tidy <- tbl_graph(nodes = nodes, edges = edges, directed = TRUE)

```

```{r}
class(routes_tidy)

```

Printing out a tbl_graph object to the console results in a drastically different output from that of an igraph object. It is an output similar to that of a normal tibble we're familiar with.

```{r}

routes_tidy


```

Visualize with ggraph.

```{r}
ggraph(routes_tidy) + 
  geom_edge_link() + 
  geom_node_point() + 
  theme_graph()


```

Well that wasn't very helpful. Let's try a different layout and parameters.

```{r}

ggraph(routes_tidy, layout = "graphopt") + 
  geom_node_point() +
  geom_edge_link(alpha = 0.8) + 
  # geom_node_text(aes(label = label), repel = TRUE) +
  theme_graph()

```

Now we'll add some color and size options.

```{r}

ggraph(routes_tidy, layout = "graphopt") + 
  geom_node_point(aes(color = group, size = .4)) +
  geom_edge_link(alpha = 0.6) + 
  # geom_node_text(aes(label = label), repel = TRUE) +
  theme_graph()

```

### Only firms that serve more than one company

To trim down our network to only the most meaningful core connections, let's filter down to the firms that only connect to at least two companies.

One way we can do this is to do some grouping/counting of our edges in much the same way we do that with a regular tibble. We'll aim to find the onces that show up at least twice, and then use that to filter the edges object itself.

```{r}

edgegroup2 <- edges %>% 
  ungroup() %>% 
  count(source) %>% 
  filter(n > 1)


edgevec2 <- edgegroup2 %>% pull(source)

edges_two_or_more <- edges %>% 
  filter(source %in% edgevec2)

head(edges_two_or_more)


```

Now let's used that same vector of values to filter the nodes.

```{r}

n1 <- nodes %>%
  select(name, group) %>% 
  filter(group == "company")

n2 <- nodes %>% 
  select(name, group) %>% 
  filter(group == "lawfirm",
         name %in% edgevec2)

nodes_two_or_more <- bind_rows(n1, n2)

nodes_two_or_more <- tibble::rowid_to_column(nodes_two_or_more, "id") %>% 
  mutate(id = id - 1) %>% 
  select(name, id, group)
  
head(nodes_two_or_more)

```

### Only firms that serve more than TWO companies (3+).

Now, let's zero in even more to the most prominent connections within the network - we'll do only those edges that have at least three connections.

```{r}

edgegroup3 <- edges %>% 
  ungroup() %>% 
  count(source) %>% 
  filter(n > 2)

edgevec3 <- edgegroup3 %>% pull(source)

edges_three_or_more <- edges %>% 
  filter(source %in% edgevec3)


n1b <- nodes %>% 
  select(name, group) %>% 
  filter(group == "company")

n2b <- nodes %>% 
  select(name, group) %>% 
  filter(group == "lawfirm",
         name %in% edgevec3)

nodes_three_or_more <- bind_rows(n1b, n2b)

nodes_three_or_more <- tibble::rowid_to_column(nodes_three_or_more, "id") %>% 
  mutate(id = id - 1) %>% 
  select(name, id, group)

```

Create the tbl_graph with our newly filtered set of values.

```{r}
# https://www.jessesadler.com/post/network-analysis-with-r/

routes_tidy_two_or_more <- tbl_graph(nodes = nodes_two_or_more, edges = edges_two_or_more, directed = FALSE)

```

```{r}

routes_tidy_two_or_more


```

Now let's visualize the newly pruned network to see what we have.

```{r}

ggraph(routes_tidy_two_or_more, layout = "graphopt") + 
  geom_node_point(aes(color = group, size = .6)) +
  geom_edge_link(alpha = 0.8) + 
  # scale_edge_width(range = c(0.2, 1)) +
  # geom_node_text(aes(label = name), repel = TRUE) +
  # labs(edge_width = "# of cases") +
  theme_graph()

```

Another layout choice.

```{r}


ggraph(routes_tidy_two_or_more, layout = "drl") + 
  geom_node_point(aes(color = group, size = .6)) +
  geom_edge_link(alpha = 0.8) + 
  # scale_edge_width(range = c(0.2, 1)) +
  # geom_node_text(aes(label = name), repel = TRUE) +
  # labs(edge_width = "# of cases") +
  theme_graph()


```

Yet another.

```{r}


ggraph(routes_tidy_two_or_more, layout = "stress") + 
  geom_node_point(aes(color = group, size = .6)) +
  geom_edge_link(alpha = 0.8) + 
  # scale_edge_width(range = c(0.2, 1)) +
  # geom_node_text(aes(label = name), repel = TRUE) +
  # labs(edge_width = "# of cases") +
  theme_graph()


```

I like this last one, so let's use that for our final version, and then add some additional options to the visualization.

```{r}

final <- ggraph(routes_tidy_two_or_more, layout = "stress") + 
  geom_node_point(aes(color = group, 
                      size = .6
                      # shape = group
                      )) +
  geom_edge_link(alpha = 0.8,
                 color = "grey66") + 
  # scale_edge_width(range = c(0.2, 1)) +
  geom_node_text(aes(filter = group=="company", label = name), 
                 family = "serif", 
                 repel = TRUE) +
  # labs(edge_width = "# of cases") +
  theme_graph()


final

```

And there we go! Nice. We've made it, and congrats on your first network visualization in R.
