---
title: "Network Analysis Visualizations"
author: "Aaron Kessler"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    code_folding: show
    df_print: kable
    toc: yes
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(kableExtra)
library(here)
library(janitor)
library(igraph)
library(tidygraph)
library(ggraph)
options(dplyr.summarise.inform = FALSE)

#load dataset
top_ten_consolidated_casecounts <- readRDS("data/table_data/top_ten_consolidated_casecounts.rds")


```

# Social Network Analysis

## Create the node and edge network datasets with igraph

Edges
```{r}

edges <- top_ten_consolidated_casecounts %>% 
  select(source = clean_firm_consolidated,
         target = parent_company, 
         weight = cases)

edges %>% head(10)

```

Nodes
```{r}

names1 <- top_ten_consolidated_casecounts %>% 
  ungroup() %>% 
  select(name = parent_company) %>% 
  distinct() %>% 
  mutate(group = "company")

names2 <- top_ten_consolidated_casecounts %>% 
  ungroup() %>% 
  select(name = clean_firm_consolidated) %>% 
  distinct() %>% 
  mutate(group = "lawfirm")

nodes <- bind_rows(names1, names2)

#add id columns with sequential values
#will zero-index the values to work with igraph more smoothly
nodes <- tibble::rowid_to_column(nodes, "id") %>% 
  mutate(id = id - 1) %>% 
  select(name, id, group)


nodes %>% 
  head(10)

```


### igraph object creation

```{r}

g <- graph_from_data_frame(d = edges, 
                           vertices = nodes, 
                           directed = FALSE)
g


```

Access nodes

```{r}

V(g)


```

Names of each node

```{r}

head(V(g)$name)


```

Edges

```{r}

E(g) # edges


```


Weights for each edge 

```{r}

E(g)$weight # weights for each edge


```

attributes of edges
```{r}

edge_attr(g) 


```


# Network visualization


## Using Tidygraph for node and edge sets

Create the tbl_graph

```{r}
# https://www.jessesadler.com/post/network-analysis-with-r/

routes_tidy <- tbl_graph(nodes = nodes, edges = edges, directed = TRUE)

```



```{r}
class(routes_tidy)


```

Printing out a tbl_graph object to the console results in a drastically different output from that of an igraph object. It is an output similar to that of a normal tibble.


```{r}

routes_tidy


```

Visualize with ggraph https://cran.r-project.org/web/packages/ggraph/index.html


```{r}
ggraph(routes_tidy) + 
  geom_edge_link() + 
  geom_node_point() + 
  theme_graph()


```



```{r}

ggraph(routes_tidy, layout = "graphopt") + 
  geom_node_point() +
  geom_edge_link(alpha = 0.8) + 
  # geom_node_text(aes(label = label), repel = TRUE) +
  theme_graph()

```



```{r}

ggraph(routes_tidy, layout = "graphopt") + 
  geom_node_point(aes(color = group, size = .4)) +
  geom_edge_link(alpha = 0.6) + 
  # geom_node_text(aes(label = label), repel = TRUE) +
  theme_graph()

```

### Only firms that serve more than one company

```{r}

edgegroup2 <- edges %>% 
  ungroup() %>% 
  count(source) %>% 
  filter(n > 1)


edgevec2 <- edgegroup2 %>% pull(source)

edges_two_or_more <- edges %>% 
  filter(source %in% edgevec2)



n1 <- nodes %>% 
  select(name, group) %>% 
  filter(group == "company")

n2 <- nodes %>% 
  select(name, group) %>% 
  filter(group == "lawfirm",
         name %in% edgevec2)

nodes_two_or_more <- bind_rows(n1, n2)

nodes_two_or_more <- tibble::rowid_to_column(nodes_two_or_more, "id") %>% 
  mutate(id = id - 1) %>% 
  select(name, id, group)

```


### Only firms that serve more than TWO companies (THREE MINIMUM)

```{r}

edgegroup3 <- edges %>% 
  ungroup() %>% 
  count(source) %>% 
  filter(n > 2)


edgevec3 <- edgegroup3 %>% pull(source)

edges_three_or_more <- edges %>% 
  filter(source %in% edgevec3)



n1b <- nodes %>% 
  select(name, group) %>% 
  filter(group == "company")

n2b <- nodes %>% 
  select(name, group) %>% 
  filter(group == "lawfirm",
         name %in% edgevec3)

nodes_three_or_more <- bind_rows(n1b, n2b)

nodes_three_or_more <- tibble::rowid_to_column(nodes_three_or_more, "id") %>% 
  mutate(id = id - 1) %>% 
  select(name, id, group)

```

Create the tbl_graph

```{r}
# https://www.jessesadler.com/post/network-analysis-with-r/

routes_tidy_two_or_more <- tbl_graph(nodes = nodes_two_or_more, edges = edges_two_or_more, directed = FALSE)

```



```{r}
class(routes_tidy_two_or_more)


```

```{r}

routes_tidy_two_or_more


```



```{r}

ggraph(routes_tidy_two_or_more, layout = "graphopt") + 
  geom_node_point(aes(color = group, size = .6)) +
  geom_edge_link(alpha = 0.8) + 
  # scale_edge_width(range = c(0.2, 1)) +
  # geom_node_text(aes(label = name), repel = TRUE) +
  # labs(edge_width = "# of cases") +
  theme_graph()



```



```{r}


ggraph(routes_tidy_two_or_more, layout = "drl") + 
  geom_node_point(aes(color = group, size = .6)) +
  geom_edge_link(alpha = 0.8) + 
  # scale_edge_width(range = c(0.2, 1)) +
  # geom_node_text(aes(label = name), repel = TRUE) +
  # labs(edge_width = "# of cases") +
  theme_graph()


```



```{r}


ggraph(routes_tidy_two_or_more, layout = "stress") + 
  geom_node_point(aes(color = group, size = .6)) +
  geom_edge_link(alpha = 0.8) + 
  # scale_edge_width(range = c(0.2, 1)) +
  # geom_node_text(aes(label = name), repel = TRUE) +
  # labs(edge_width = "# of cases") +
  theme_graph()


```



```{r}

final <- ggraph(routes_tidy_two_or_more, layout = "stress") + 
  geom_node_point(aes(color = group, 
                      size = .6
                      # shape = group
                      )) +
  geom_edge_link(alpha = 0.8,
                 color = "grey66") + 
  # scale_edge_width(range = c(0.2, 1)) +
  geom_node_text(aes(filter = group=="company", label = name), 
                 family = "serif", 
                 repel = TRUE) +
  # labs(edge_width = "# of cases") +
  theme_graph()


final

```
